### 2022-10-22 dissertation/chapters/01-introduction.tex
🚧 (chapter3): WIP - Save point

\begin{itemize}
      \item \textbf{Section \ref{sec:introduction:summary_research_domain}} provides the reader with brief summary of the research domain.

      \item \textbf{Section \ref{sec:introduction:problem}} outlines the research problem being addressed.

      \item \textbf{Section \ref{sec:introduction:motivation}} provides a motivation for this research.

      \item \textbf{Section \ref{sec:introduction:objectives}} presents the reader with the research objectives and outlines the goals of this dissertation.

      \item \textbf{Section \ref{sec:introduction:contributions}} outlines the contributions of this research to the field.

      \item \textbf{Section \ref{sec:introduction:outline}} provides a full dissertation outline.
\end{itemize}

@@

In Section \ref{sec:introduction:problem} above, it is mentioned that finding the best heuristic to train \acp{FFNN} is a timely and tedious process. A modern approach is to use \acp{HH} to automate the process of selecting the best heuristic when applied to some optimisation problem. The best heuristic might not be a single heuristic, but rather a hybridisation of heuristics \cite{ref:pillay:2015}.

@@

\begin{itemize}
      \item \textbf{Appendix~\ref{app:datasets}} provides details on the datasets used for the empirical analysis.

@@

      \item \textbf{Appendix~\ref{app:derived_publications}} lists the publications derived from this work.

      \item \textbf{Appendix~\ref{app:symbols}} lists and defines the mathematical symbols used in this work, categorised according to the relevant chapter in which they appear.
\end{itemize}

>>>>

\begin{itemize}
      \item \textbf{Section~\ref{sec:introduction:summary_research_domain}} provides the reader with brief summary of the research domain.

      \item \textbf{Section~\ref{sec:introduction:problem}} outlines the research problem being addressed.

      \item \textbf{Section~\ref{sec:introduction:motivation}} provides a motivation for this research.

      \item \textbf{Section~\ref{sec:introduction:objectives}} presents the reader with the research objectives and outlines the goals of this dissertation.

      \item \textbf{Section~\ref{sec:introduction:contributions}} outlines the contributions of this research to the field.

      \item \textbf{Section~\ref{sec:introduction:outline}} provides a full dissertation outline.
\end{itemize}

@@

In Section~\ref{sec:introduction:problem} above, it is mentioned that finding the best heuristic to train \acp{FFNN} is a timely and tedious process. A modern approach is to use \acp{HH} to automate the process of selecting the best heuristic when applied to some optimisation problem. The best heuristic might not be a single heuristic, but rather a hybridisation of heuristics \cite{ref:pillay:2015}.

@@

\begin{itemize}
      \item \textbf{Appendix~\ref{app:backpropagation}} provides a full breakdown of the \acf{BP} algorithm as presented by \citeauthor{ref:engelbrecht:2007} \cite{ref:engelbrecht:2007}.

      \item \textbf{Appendix~\ref{app:symbols}} lists and defines the mathematical symbols used in this work, categorised according to the relevant chapter in which they appear.

      \item \textbf{Appendix~\ref{app:datasets}} provides details on the datasets used for the empirical analysis.

@@

      \item \textbf{Appendix~\ref{app:derived_publications}} lists the publications derived from this work.
\end{itemize}

### 2022-10-22 dissertation/chapters/02-artificial-neural-networks.tex
🚧 (chapter3): WIP - Save point

\begin{itemize}
    \item \textbf{Section \ref{sec:anns:bn}} gives background information on the \acs{BN}.

    \item \textbf{Section \ref{sec:anns:an}} introduces the \acs{AN}. Brief discussions follow on input, \index{weights}weights and \index{biases}biases, \index{net input signal}net input signal, \index{activation function}activation functions, and output.

    \item \textbf{Section \ref{sec:anns:ann}} introduces the \acs{ANN}. Brief discussions follow on \acs{ANN} architecture, topology, and \acp{FFNN}.

    \item \textbf{Section \ref{sec:anns:training}} provides details on the training process, \index {supervised learning}supervised learning, training sets, stopping conditions, performance measures and \index{error function}error functions.

    \item \textbf{Section \ref{sec:anns:summary}} provides a brief summary of the chapter.
\end{itemize}

@@

\noindent
The biological neural systems are made up of microscopic nerve cells called neurons \cite{ref:jain:1996}. Figure \ref{fig:biological_neuron} illustrates a single \acs{BN}. The main components of the \acs{BN} include the cell body, \index{dendrites}\textit{dendrites} and the \index{axon}\textit{axon}. \Acp{NN} are formed through the connections between the \index{axon}axons and \index{dendrites}dendrites of various neurons. This is known as \index{synaptogenesis}\textit{synaptogenesis} \cite{ref:huttenlocher:1997}. Such a connection is referred to as a \index{synapse}\textit{synapse}. Communication takes place, through the synapse, by electro-chemical pulse and is often referred to as an \textit{activation} or \textit{action potential}.  Communication signals propagate from the \index{dendrites}dendrites, through the cell body to the axon of a neuron, provoking a signal in the post-synaptic neuron \cite{ref:engelbrecht:2007}. The greater the connection between two neurons, the stronger the communication.  Kennedy \cite{ref:kennedy:2016} defines stronger \index{synapse}synapses as ones that contribute more depolarisation to the neural membrane upon activation than weaker ones. Stronger \index{synapse}synapses have a higher probability of generating an action potential in the post-synaptic neuron. During activation, the pre-synaptic neuron release neurotransmitters that bind to the post-synaptic neuron \cite{ref:khanacademy:synapse}. The frequent release of these molecules cause the \index{synapse}synapse to grow. Connections that grow over time yield stronger signals (learning), while connections that are weak propagate low intensity signals and vanish over time (forgetting). The ability of synapses to strengthen and weaken over time is known as \index{synaptic plasticity}\textit{synaptic plasticity} \cite{ref:huttenlocher:1997}.

@@

The \acs{AN} implements various components and is illustrated in Figure \ref{fig:artificial_neuron}.

@@

\begin{itemize}
    \item \textbf{Input:} The input models the activations received from the pre-synaptic neuron or from some environment sensor. Input is represented by the input vector $\vec{x}$, in Figure \ref{fig:artificial_neuron}.

    \item \textbf{Weights:} The weights model the \index{synapse}synapses and connection strengths. Weights are represented by the weight vector $\vec{v}$, in Figure \ref{fig:artificial_neuron}.

    \item \index{net input signal}\textbf{Net input signal:} The \index{net input signal}net input signal models the net resulting activations from all connected pre-synaptic neurons or environment sensors. The \index{net input signal}net input signal is represented by $net$, in Figure \ref{fig:artificial_neuron}.

    \item \textbf{Biases:} The biases model a mechanism introduced to influence the strength of the activation (output signal) of the \acs{AN}. The bias is represented by $\theta$, in Figure \ref{fig:artificial_neuron}.

    \item \index{activation function}\textbf{Activation function:} The activation function models the action potential of the \acs{BN} and is based on the net input signal. The \index{activation function}activation function is represented by $f$, in Figure \ref{fig:artificial_neuron}.

    \item \textbf{Output:} The output models the activation by the post-synaptic neuron. Output is represented by the output vector $\vec{y}$, in Figure \ref{fig:artificial_neuron}.
\end{itemize}

@@

\noindent
By introducing a parameter $\alpha > 0$, negative \index{net input signal}net input signals will still yield non-zero activations, resulting in non-zero gradients and avoiding gradient saturation. Non-zero gradients are required by some heuristics such as \acs{SGD} in order to be able to effectively train \acp{ANN} \cite{ref:hanin:2018}. An illustration of \acs{LReLU} with various values for $\alpha$ and $\theta = 0$ is given in Figure \ref{fig:anns:activation_functions:leaky_relu}.

@@

\noindent
In Equation~\eqref{eq:sigmoid} above, $\lambda$ is a control parameter, which controls the steepness of the sigmoid activation function and is usually set to $\lambda = 1$. An illustration of the \index{sigmoid}sigmoid \index{activation function}activation function with various values for $\lambda$ and $\theta = 0$ is given in Figure \ref{fig:anns:activation_functions:sigmoid}.

@@

\noindent
An illustration of the \index{hyperbolic tangent}hyperbolic tangent \index{activation function}activation function with various values for $\lambda$ and $\theta = 0$ is given in Figure \ref{fig:anns:activation_functions:hyperbolic_tangent}.

@@

\noindent
Figure \ref{fig:anns:activation_functions:softmax_argmax} illustrates the comparison of transformations of the output vector $\vec{y}$, where the \index{sigmoid}sigmoid, the \index{softmax}softmax and the \index{argmax}argmax \index{activation function}activation functions are used.

@@

\acp{FFNN} were the first and simplest type of \acp{ANN} developed \cite{ref:schmidhuber:2015} and implement input, hidden and output layers by arranging them in sequential order. Furthermore, \acp{FFNN} implement \index{fully connected topology}fully connected topologies. In \acp{FFNN}, information moves forward, in one direction, from the input nodes, through the hidden nodes and finally to the output nodes. Depending on the optimisation algorithm used, error correction information can be ``backpropagated'' through the network. An illustration of a \acs{FFNN} is given in Figure \ref{fig:ffnn} below.

@@

\noindent
In Figure \ref{fig:ffnn}, $x_i$ refers to the $i$-th dimension in the input vector $\vec{x}$, $h_j$ refers to the $j$-th dimension in the hidden layer, $y_k$ refers to the $k$-th dimension in the output vector $\vec{y}$, $v_{i,j}$ refers to the weight associated with input node $x_i$ and the hidden node $h_j$, and $w_{j,k}$ refers to the weight associated with hidden node $h_j$ and the output node $y_k$.

@@

Finding the optimal weights that produce the best performance on a given task is an optimisation problem. The optimisation algorithm used to find the optimal weights is referred to as a \index{heuristic}\textit{heuristic}. \index{heuristic}Heuristics search for possible solutions in the solution-space and make use of information from the search-space to guide to process. Heuristics are discussed in detail in Chapter \ref{chap:heuristics}.

@@

\subsubsection{Categorical Cross-Entropy}
\label{sec:anns:training:error_functions:cat_xe}

@@

\section{Summary}
\label{sec:anns:summary}

>>>>

\begin{itemize}
    \item \textbf{Section~\ref{sec:anns:bn}} gives background information on the \acs{BN}.

    \item \textbf{Section~\ref{sec:anns:an}} introduces the \acs{AN}. Brief discussions follow on input, \index{weights}weights and \index{biases}biases, \index{net input signal}net input signal, \index{activation function}activation functions, and output.

    \item \textbf{Section~\ref{sec:anns:ann}} introduces the \acs{ANN}. Brief discussions follow on \acs{ANN} architecture, topology, and \acp{FFNN}.

    \item \textbf{Section~\ref{sec:anns:training}} provides details on the training process, \index {supervised learning}supervised learning, training sets, stopping conditions, performance measures and \index{error function}error functions.

    \item \textbf{Section~\ref{sec:anns:summary}} provides a brief summary of the chapter.
\end{itemize}

@@

\noindent
The biological neural systems are made up of microscopic nerve cells called neurons \cite{ref:jain:1996}. Figure~\ref{fig:biological_neuron} illustrates a single \acs{BN}. The main components of the \acs{BN} include the cell body, \index{dendrites}\textit{dendrites} and the \index{axon}\textit{axon}. \Acp{NN} are formed through the connections between the \index{axon}axons and \index{dendrites}dendrites of various neurons. This is known as \index{synaptogenesis}\textit{synaptogenesis} \cite{ref:huttenlocher:1997}. Such a connection is referred to as a \index{synapse}\textit{synapse}. Communication takes place, through the synapse, by electro-chemical pulse and is often referred to as an \textit{activation} or \textit{action potential}.  Communication signals propagate from the \index{dendrites}dendrites, through the cell body to the axon of a neuron, provoking a signal in the post-synaptic neuron \cite{ref:engelbrecht:2007}. The greater the connection between two neurons, the stronger the communication.  Kennedy \cite{ref:kennedy:2016} defines stronger \index{synapse}synapses as ones that contribute more depolarisation to the neural membrane upon activation than weaker ones. Stronger \index{synapse}synapses have a higher probability of generating an action potential in the post-synaptic neuron. During activation, the pre-synaptic neuron release neurotransmitters that bind to the post-synaptic neuron \cite{ref:khanacademy:synapse}. The frequent release of these molecules cause the \index{synapse}synapse to grow. Connections that grow over time yield stronger signals (learning), while connections that are weak propagate low intensity signals and vanish over time (forgetting). The ability of synapses to strengthen and weaken over time is known as \index{synaptic plasticity}\textit{synaptic plasticity} \cite{ref:huttenlocher:1997}.

@@

The \acs{AN} implements various components and is illustrated in Figure~\ref{fig:artificial_neuron}.

@@

\begin{itemize}
    \item \textbf{Input:} The input models the activations received from the pre-synaptic neuron or from some environment sensor. Input is represented by the input vector $\vec{x}$, in Figure~\ref{fig:artificial_neuron}.

    \item \textbf{Weights:} The weights model the \index{synapse}synapses and connection strengths. Weights are represented by the weight vector $\vec{v}$, in Figure~\ref{fig:artificial_neuron}.

    \item \index{net input signal}\textbf{Net input signal:} The \index{net input signal}net input signal models the net resulting activations from all connected pre-synaptic neurons or environment sensors. The \index{net input signal}net input signal is represented by $net$, in Figure~\ref{fig:artificial_neuron}.

    \item \textbf{Biases:} The biases model a mechanism introduced to influence the strength of the activation (output signal) of the \acs{AN}. The bias is represented by $\theta$, in Figure~\ref{fig:artificial_neuron}.

    \item \index{activation function}\textbf{Activation function:} The activation function models the action potential of the \acs{BN} and is based on the net input signal. The \index{activation function}activation function is represented by $f$, in Figure~\ref{fig:artificial_neuron}.

    \item \textbf{Output:} The output models the activation by the post-synaptic neuron. Output is represented by the output vector $\vec{y}$, in Figure~\ref{fig:artificial_neuron}.
\end{itemize}

@@

\noindent
By introducing a parameter $\alpha > 0$, negative \index{net input signal}net input signals will still yield non-zero activations, resulting in non-zero gradients and avoiding gradient saturation. Non-zero gradients are required by some heuristics such as \acs{SGD} in order to be able to effectively train \acp{ANN} \cite{ref:hanin:2018}. An illustration of \acs{LReLU} with various values for $\alpha$ and $\theta = 0$ is given in Figure~\ref{fig:anns:activation_functions:leaky_relu}.

@@

\noindent
In Equation~\eqref{eq:sigmoid} above, $\lambda$ is a control parameter, which controls the steepness of the sigmoid activation function and is usually set to $\lambda = 1$. An illustration of the \index{sigmoid}sigmoid \index{activation function}activation function with various values for $\lambda$ and $\theta = 0$ is given in Figure~\ref{fig:anns:activation_functions:sigmoid}.

@@

\noindent
An illustration of the \index{hyperbolic tangent}hyperbolic tangent \index{activation function}activation function with various values for $\lambda$ and $\theta = 0$ is given in Figure~\ref{fig:anns:activation_functions:hyperbolic_tangent}.

@@

\noindent
Figure~\ref{fig:anns:activation_functions:softmax_argmax} illustrates the comparison of transformations of the output vector $\vec{y}$, where the \index{sigmoid}sigmoid, the \index{softmax}softmax and the \index{argmax}argmax \index{activation function}activation functions are used.

@@

\acp{FFNN} were the first and simplest type of \acp{ANN} developed \cite{ref:schmidhuber:2015} and implement input, hidden and output layers by arranging them in sequential order. Furthermore, \acp{FFNN} implement \index{fully connected topology}fully connected topologies. In \acp{FFNN}, information moves forward, in one direction, from the input nodes, through the hidden nodes and finally to the output nodes. Depending on the optimisation algorithm used, error correction information can be ``backpropagated'' through the network. An illustration of a \acs{FFNN} is given in Figure~\ref{fig:ffnn} below.

@@

\noindent
In Figure~\ref{fig:ffnn}, $x_i$ refers to the $i$-th dimension in the input vector $\vec{x}$, $h_j$ refers to the $j$-th dimension in the hidden layer, $y_k$ refers to the $k$-th dimension in the output vector $\vec{y}$, $v_{i,j}$ refers to the weight associated with input node $x_i$ and the hidden node $h_j$, and $w_{j,k}$ refers to the weight associated with hidden node $h_j$ and the output node $y_k$.

@@

Finding the optimal weights that produce the best performance on a given task is an optimisation problem. The optimisation algorithm used to find the optimal weights is referred to as a \index{heuristic}\textit{heuristic}. \index{heuristic}Heuristics search for possible solutions in the solution-space and make use of information from the search-space to guide to process. Heuristics are discussed in detail in Chapter~\ref{chap:heuristics}.

@@

\subsubsection{Categorical Cross-Entropy}\label{sec:anns:training:error_functions:cat_xe}

@@

\section{Summary}\label{sec:anns:summary}
